### 2014-12-03 doc/GMP2015/s00_abstract.tex

\begin{abstract}
Super awesome abstract.
\end{abstract}