### 2014-12-18 doc/GMP2015/paper.tex
Added another observation about algebraic complexity

%
% include the class file for GMP 2015
%
\documentclass[submission]{gmp2015}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{tikz}
\usetikzlibrary{shapes,snakes}
\newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}

\usepackage{ulem}

\newcommand{\todo}[1]{\textsc{{\color{red}#1}}}

\usepackage{caption}
\usepackage{subcaption}
\usepackage{floatrow}
% \usepackage{minipage}

\newtheorem{observation}[theorem]{\sffamily Observation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%
% put the title of your paper here
%
\title{Designing Qusecs Optimally:
Quasi-uniform or Self-similar Layered Materials as
Recursively Decomposed Solutions of Geometric Constraint
Systems}

%
% put your submission number here
% this number will appear instead of authors and affiliations if the class option "submission" is active
%
\SubNumber{110}

%
% put the author names here
% use the second argument as a reference to the list of affiliations
% no authors and affiliations will appear if the class option "submission" is active
%
\author{Troy Baker}{1}
\author{Meera Sitharam}{1}
\author{Menghan Wang}{1}
\author{Joel Willoughby}{1}

%
% put the affiliations of the authors here
%
\affiliation{1}{Department of Computer and Information Science and Engineering, University of Florida, USA}

%
% the rest is as usual
%

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{s00_abstract}

\input{s01_introduction}

\input{s02_preliminaries}

\input{s03_drp}

\input{s04_recombination}

\input{s09_problems}

\input{s07_problempinning}

\input{s08_pinnedline}

\input{s06_conclusion}

% \begin{figure}\centering
%   \includegraphics[scale=.5]{example.eps}
%   \caption{Example of a normal figure that spans onse column.}
%   \label{fig:square}
% \end{figure}

% \begin{figure*}\centering
%     \includegraphics{example.eps}
%     \caption{Example of a figure that spans both columns.}
% \end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrv}
\bibliography{paper}

\clearpage
\input{appendix_a}

\end{document}