### 2014-12-18 doc/GMP2015/s09_conclusion.tex
Started a conclusion

\section{Conclusion}
We have worked to navigate the complexity barriers faced by the general DR-plan and recombination problem. To this end we have defined a new canonical DR-plan and given an effiecient ($O(n^2)$) algorithm to find the DR-plan for specific types of 2D qusecs.

We have also described methods of recombining the DR-plan to get a realization different from the normal solving of a system of equations. These results rely on the results obtained in the Cayley Configuration space of certain graphs.

We then related these theoretical results to real-world situations in which they can be applied including Examples 1-5 in Section \ref{sec:intro}.

Throughout the paper, we raised some open questions and future areas of research. These include: completion of body-hyper-pin qusecs with more than 2-dof, \todo{Add a general synopsis of other questions}