### 2014-12-19 doc/GMP2015/s09_conclusion.tex

\section{Conclusion}
\label{sec:conclusion}
% \bibliography{paper}

We have elucidated the complexity barriers, for example over-constraints, faced by the general optimal DR-plan and recombination problem. When there are no overconstraints, which is the case for 2D qusecs whose realizations are many common types of layered materials, we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property given an efficient ($O(n^2)$) algorithm to find a DR-plan that satisfies other desirable characteristics.

We have also described novel methods of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise non-decomposable system that recombines the solved child systems at a node of a DR-plan. These results rely on a recent theory of Convex Cayley Configuration spaces.

We then modeled specific layered materials using extensions of the above theoretical results including the motivating Examples 1-5 in the introduction.

In Appendix B [\ref{sec:appendix:b}] we provide open problems for future work.