### 2014-12-20 doc/GMP2015/s09_conclusion.tex

\section{Conclusion}
% \bibliography{paper}

We have clarified the main source of complexity for the optimal DR-plan and recombination problems. For the former problem, when there are no overconstraints, (as is the case for 2D qusecs whose realizations are many common types of layered materials), we defined a canonical DR-plan and showed that such canonical DR-plans are guaranteed to be optimal, a strong Church-Rosser property giving an efficient ($O(n^2)$) algorithm to find an optimal  DR-plan that satisfies other desirable characteristics.

We have also described a novel method of efficiently realizing a 2D qusecs from the optimal DR-plan by modifying the otherwise nondecomposable system that recombines the solved child systems at a node of a DR-plan These results rely on a recent theory of Convex Cayley Configuration spaces. Relationships and reductions between these and previously studied problems were formally clarified.

We then modeled specific layered materials using extensions of the above theoretical results including the motivating Examples 1-5 in the introduction.

In Appendix B [\ref{sec:appendix:b}] we provide open problems for future work.