### 2014-12-20 doc/GMP2015/paper.tex

%
% include the class file for GMP 2015
%
\documentclass[submission]{gmp2015}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{tikz}
\usetikzlibrary{shapes}
% \usepackage{ulem} % This was breaking emph!
% \usepackage[normalem]{ulem}
\usepackage{soul}
\usepackage{caption}
\usepackage{subcaption}
% \usepackage{floatrow}

% \usepackage{atbeginend}

\newtheorem{observation}[theorem]{\sffamily Observation}
% \newtheorem{openproblem}[theorem]{\sffamily Open Problem}

\theoremstyle{definition}
\newtheorem{openproblem}[theorem]{\sffamily Open Problem}
% \theoremstyle{theorem}
% \newenvironment{openproblem}{
%     % \begin{samepage}
%     \begin{trivlist}
%         \item[\hskip \labelsep \textbf{Open Problem:}]
% }
% {
%     \end{trivlist}
%     % \end{samepage}
% }

\newcommand{\dfn}[1]{\textit{#1}} % definition of a word
\newcommand{\vemph}[1]{\textsl{#1}} % verbal emphasis
% \newcommand{\header}[1]{\underbar{\textsl{#1}}}
\newcommand{\header}[1]{\ul{\textsl{#1}}}
\newcommand{\problemstatement}[1]{\medskip\noindent\textbf{#1}}
\newcommand{\myexample}{\medskip\noindent\textbf{Example.} }

\newcommand{\note}{\textbf{Note: }}

\newcommand{\todo}[1]{\textsc{{\color{red}#1}}}
\renewcommand{\todo}[1]{}
\newcommand{\uncited}{\textsc{{\bf\color{red}[??]}}}

\newcommand{\tedge}[5]{\draw[#3] (#1)-- node[e, #5] (e#4) {#4} (#2)}

\newcommand{\drp}[1]{\text{DRP}(#1)}
\newcommand{\comdrp}[1]{\text{ComDRP}(#1)}
\newcommand{\optdrp}[1]{\text{OptDRP}(#1)}
\newcommand{\candrp}[1]{\text{CanDRP}(#1)}

\newcommand{\candrpcomplexity}{$O(n^3)$}
\newcommand{\candrpcomplexityv}{$O(|V|^3)$}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%
% put the title of your paper here
%
\title{Designing Qusecs Optimally:
Quasi-uniform or Self-similar Layered Materials as
Recursively Decomposed Solutions of Geometric Constraint
Systems}

%
% put your submission number here
% this number will appear instead of authors and affiliations if the class option "submission" is active
%
\SubNumber{110}

%
% put the author names here
% use the second argument as a reference to the list of affiliations
% no authors and affiliations will appear if the class option "submission" is active
%
\author{Troy Baker}{1}
\author{Meera Sitharam}{1}
\author{Menghan Wang}{1}
\author{Joel Willoughby}{1}

%
% put the affiliations of the authors here
%
\affiliation{1}{Department of Computer and Information Science and Engineering, University of Florida, USA}

\linespread{0.9}
% \columnsep{0.95}

% \renewenvironment{itemize}{\begin{itemize}
% \addtolength{\topsep}{-\baselineskip}
% \addtolength{\itemsep}{-\baselineskip}
% \end{itemize}}
\let\tempone\itemize
\let\temptwo\enditemize
\renewenvironment{itemize}{
\tempone
    \addtolength{\topsep}{-\baselineskip}
    \addtolength{\partopsep}{-\baselineskip}
    \addtolength{\itemsep}{-0.25\baselineskip}
}{\temptwo}
\let\tempone\enumerate
\let\temptwo\endenumerate
\renewenvironment{enumerate}{
\tempone
    \addtolength{\topsep}{-\baselineskip}
    \addtolength{\partopsep}{-\baselineskip}
    \addtolength{\itemsep}{-0.25\baselineskip}
}{\temptwo}

% \setlist[itemize]{noitemsep, topsep=0pt}

%
% the rest is as usual
%

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{s00_abstract}

\input{s01_introduction}

\input{s02_preliminaries}

\input{s03_drp}

\input{s04_recombination}

\input{s05_problems}

\input{s07_problempinning}

\input{s08_pinnedline}

\input{s09_conclusion}

% \begin{figure}\centering
%   \includegraphics[scale=.5]{example.eps}
%   \caption{Example of a normal figure that spans onse column.}
%   \label{fig:square}
% \end{figure}

% \begin{figure*}\centering
%     \includegraphics{example.eps}
%     \caption{Example of a figure that spans both columns.}
% \end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrv}
\bibliography{paper}
% {\scriptsize \bibliography{paper}}

\clearpage

\linespread{1}
\input{appendix_a}
\input{appendix_b}

\end{document}