### 2015-02-16 doc/GMP2015/revision_1/s04_recombination.tex

\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
\label{sec:recomb}
%
In this section, we consider the \dfn{optimal recombination} problem of combining specific solutions of subsystems in a DR-plan into a solution of their parent system $I$ (without loss of generality, at the top level of the DR-plan). In the case of isostatic qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we seek \vemph{solution(s) (among a finite large number of solutions) with a specific orientation or chirality}. In the case of underconstrained qusecs the subsystems are the multiple roots of the DR-plan, the parent system $I$ is underconstrained, and we typically seek an efficient algorithmic description of \vemph{connected component(s) of solutions with a specific orientation or chirality}.

The main barrier in recombination when given an optimal DR-plan (of smallest possible size or max fan-in),  is that the number of children (resp. number of roots)---and correspondingly the  size and complexity of the (indecomposable) algebraic system $I$ to be solved---could be arbitrarily large as a function of the size of the input system. This difficulty can persist even after optimal parametrization of the indecomposable system $I$ as in \cite{sitharam2010optimized} to minimize its algebraic complexity.

In the following, we formulate the problem of \dfn{optimal modification}, of the indecomposable algebraic system $I$ into a decomposable system with a small DR-plan (low algebraic complexity). Leveraging recent results on \dfn{Cayley configuration spaces}, our approach to the optimal modification problem achieves the following:

\medskip\noindent
\header{(a) Small DR-plan.}
We obtain a  parameterized family of systems $I_{\lambda_F}$---one for each value $\lambda_F$ for the parameters $F$,  all of which have small DR-plans. Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can potentially be solved or realized easily once the orientation type of the solution is known  (when the DR-plan size is small enough).

\medskip\noindent
\header{(b) Solution preservation.}
Moreover, the union of solution spaces of the systems in the family $I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions.

\medskip\noindent
\header{(c) Efficient search.}
Finally, this union can be projected into the so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is convex or otherwise easy to traverse in order to search for $I$'s solution (or connected component) of the desired orientation type. For the case when the modification is bounded, this approach provides an efficient algorithm for recombination. We first define the decision version of the problem of optimal modification for decomposition. The standard optimization versions are straightforward.

\problemstatement{Optimal Modification for Decomposition (OMD) Problem.}
Given a generically independent graph $G = (V,E)$ with no non-trivial proper isostatic subgraph (indecomposable) and 2 constants $k$ and $s$, does there exist a set of at most $k$  edges $E_1$ and a set of non-edges $F$ such that the modified graph $H = (V, E\setminus E_1 \cup F)$ has a DR-plan of size at most $s$?  The OMD$_k$ problem is  OMD where $k$ is a fixed bound (not part of the input).   We say that such a tuple $(G,s)$ is a member of the set OMD$_k$. We loosely refer to graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs with appropriately small $s$.

It is immediately clear that indecomposable graphs $G$ that belong in OMD$_k$ for small $k$ and $s$  lend themselves to modification  into decomposable graphs satisfying Criteria (a) and (b) above. However, it is not clear how Criterion (c) is met by OMD graphs.
%
\subsection{Using Convex Cayley Configuration Spaces}
\label{sec:2-tree-reduction}
%
Next we provide the necessary background to describe a specific approach for achieving the requirements (a)--(c) mentioned above, by restricting the class of reduced graphs $G' = G\setminus E_1$ and their isostatic completions $H$ in the above definition of the OMD problem, and using a key theorem of Convex Cayley configuration spaces \cite{sitharam2010convex}. This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley parameters), such that the set of vectors $\lambda_F$ of  attainable lengths of the non-edges $F$ is always convex for any given lengths $\delta$ for the edges of $H$ (i.e, over all the realizations of the bar-joint constraint system or linkage $(H,\delta)$ in 2 dimensions). This set is called the (2-dimensional) \dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as a ``projection'' of the cartesian realization space of $(H,\delta)$ on the Cayley parameters $F$. Such graphs $H$ are said to have \dfn{convexifiable Cayley configuration spaces with parameters $F$} (short: \dfn{convexifiable}).

To state the theorem, we first have to define the notion of \dfn{2-sums} and \dfn{2-trees}. Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$ respectively.% Let $H_1=(V_1,E_1)$ and $H_2=(V_2,E_2)$ be two graphs on disjoint sets of vertices that contain the edges $(u,v)$ and $(w,x)$ respectively.
A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$ obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$ and $v=w$. In this case, $H_1$ and $H_2$ are called \dfn{2-sum components} of $H$. A \dfn{minimal 2-sum component} of $H$ is  one that cannot be further split into 2-sum components. A \dfn{2-tree} is recursively obtained by taking a 2-sum of 2-trees, with the base case of a 2-tree being a triangle. A \dfn{partial 2-tree} is a 2-tree minus some edges. Partial 2-trees have an alternate characterization as the graphs that avoid $K_4$ minors, and are also called series parallel graphs.

\begin{theorem}\label{theorem:convexcayley}
    \cite{sitharam2010convex} $H$ has a convexifiable Cayley configuration space  with parameters $F$ if and only if for each $f\in F$  all the minimal 2-sum components of $H\cup F$ that contain both endpoints of $f$ are partial 2-trees. The Cayley configuration space $\Phi_F(H,\delta)$ of a bar-joint system or linkage $(H,\delta)$ is a convex polytope. When $H\cup F$ is a 2-tree, the bounding hyperplanes of this polytope are triangle inequalities relating the lengths of edges of the triangles in $H\cup F$.
\end{theorem}

The idea of our approach to achieve the criteria (a)--(c) begins with the following simple but useful theorem.

\begin{theorem}\label{theorem:omdk}
    Given an indecomposable graph $G$, let $G'$ be a spanning partial 2-tree subgraph in $G$ with $k$ fewer edges than $G$. Then $(G,2)$ belongs in the set OMD$_k$.
\end{theorem}

% \begin{proof}
% The proof follows from the fact that 2-trees are well decomposable and
% have simple DR-plans of size 2. We know that $G$ can be reduced by
% removing $k$ edges to create a partial-2-tree $G'$ which can then be
% completed to an (isostatic) 2-tree by adding some set of non-edges
% $F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2,
% proving the theorem.
% \end{proof}

We refer to such graphs $G$ in short as \dfn{$k$-approximately convexifiable}, where the reduced graphs $G'$ and isostatic completions $H$ are convexifiable. As observed earlier, since graphs such as $G$ are in OMD$_k$, Criteria (a) and (b) are automatically met for small enough $k$. Criterion (c) is addressed as described in the following efficient search procedure which clarifies the dependence of the complexity on the number and ranges  of Cayley parameters $F$.

\begin{theorem}\label{theorem:criterionc}
    [Efficient search]
    For an indecomposable, $k$-approximately convexifiable graph $G = (V,E)$, let $G' = (V,E' =E\setminus D)$ be a spanning partial 2-tree subgraph where $|D| \le  k$. Let  $F$ be a set of non-edges of $G$ such that $H = (V, E'\cup F)$ is a 2-tree. Each solution $p$ (or connected component of a solution space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be found in time $O(\log(W))$ where $W$ is the number of cells of desired accuracy (discrete volume) of the convex polytope $\Phi_F(G',\delta_E')$. The (discrete) volume $W$ is exponential in $|F|$ and polynomial in the (discrete range) of the parameters in $F$.
\end{theorem}

% \begin{proof}
%     The Cartesian realization space of $(H,\left<\delta_{E'},
%     \lambda_F\right>)$ is computed easily with a DR-plan of size 2,
%     and is the union of $2^t$ solutions (modulo orientation preserving
%     isometries) each with a distinct orientation type, where $t$ is
%     the number of triangles in the 2-tree $H$; here $\delta_{S}$ is
%     the restriction of the length vector $\delta$ to the edges in $S$.
%     A desired solution $p$ (or connected component of a solution
%     space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
%     found by a subdivided binary search   of the cartesian realization
%     space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation
%     type $\sigma_p$ - as $\lambda_F$ ranges over the discretized
%     convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes
%     described in Theorem \ref{theorem:convexcayley}. A solution $p$
%     is found  when the lengths for nonedges in $D$ match $\delta_D$.
% \end{proof}
%
% \begin{figure*}\centering
% \begin{subfigure}{.2\linewidth}\centering
%     \input{img/tikz/omd_k33_example_new-0.tikz}
%     \caption{}\label{fig:omd_k33_example:a}
% \end{subfigure}\hfill
% \begin{subfigure}{.15\linewidth}\centering
%     \input{img/tikz/omd_k33_example_new-1.tikz}
%     \caption{}\label{fig:omd_k33_example:b}
% \end{subfigure}
% \begin{subfigure}{.15\linewidth}\centering
%     \input{img/tikz/omd_k33_example_new-2.tikz}
%     \caption{}\label{fig:omd_k33_example:c}
% \end{subfigure}\hfill
% \begin{subfigure}{.15\linewidth}\centering
%     \input{img/tikz/omd_k33_example-1.tikz}
%     \caption{}\label{fig:omd_k33_example:d}
% \end{subfigure}
% \begin{subfigure}{.15\linewidth}\centering
%     \input{img/tikz/omd_k33_example-2.tikz}
%     \caption{}\label{fig:omd_k33_example:e}
% \end{subfigure}

% \caption{
% (\ref{fig:omd_k33_example:a}) The $K_{3,3}$ with two labeled edges, $e_1$ and $e_2$. (\ref{fig:omd_k33_example:b}) The $K_{3,3}$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree. (\ref{fig:omd_k33_example:c}) The $K_{3,3}$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K_{3,3}$ is at least $OMD_2$. (\ref{fig:omd_k33_example:d}) The $K_{3,3}$ with only $e_2$ removed (dashed line). (\ref{fig:omd_k33_example:e}) The $K_{3,3}$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K_{3,3}$ is $OMD_1$.
% }
% \label{fig:omd_k33_example}
% \end{figure*}

\ClearMyMinHeight
\SetMyMinHeight{.2}{img/epsfromtikz/omd_k33_example_new-0}
\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example_new-1}
\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example_new-2}
\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example-1}
\SetMyMinHeight{.15}{img/epsfromtikz/omd_k33_example-2}

\begin{figure*}\centering
    \begin{subfigure}{.2\linewidth}\centering
        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example_new-0}
        \caption{}\label{fig:omd_k33_example:a}
    \end{subfigure}\hfill
    \begin{subfigure}{.15\linewidth}\centering
        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example_new-1}
        \caption{}\label{fig:omd_k33_example:b}
    \end{subfigure}
    \begin{subfigure}{.15\linewidth}\centering
        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example_new-2}
        \caption{}\label{fig:omd_k33_example:c}
    \end{subfigure}\hfill
    \begin{subfigure}{.15\linewidth}\centering
        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example-1}
        \caption{}\label{fig:omd_k33_example:d}
    \end{subfigure}
    \begin{subfigure}{.15\linewidth}\centering
        \includegraphics[height=\myMinHeight]{img/epsfromtikz/omd_k33_example-2}
        \caption{}\label{fig:omd_k33_example:e}
    \end{subfigure}

    \caption{
    (\ref{fig:omd_k33_example:a}) The $K_{3,3}$ with two labeled edges, $e_1$ and $e_2$. (\ref{fig:omd_k33_example:b}) The $K_{3,3}$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree. (\ref{fig:omd_k33_example:c}) The $K_{3,3}$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K_{3,3}$ is at least $OMD_2$. (\ref{fig:omd_k33_example:d}) The $K_{3,3}$ with only $e_2$ removed (dashed line). (\ref{fig:omd_k33_example:e}) The $K_{3,3}$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K_{3,3}$ is $OMD_1$.
    }
    \label{fig:omd_k33_example}
\end{figure*}

% \noindent
% \note A major advantage of the convex Cayley method is that it is completely unaffected when $\delta$ are intervals rather than exact values~~\cite{sitharam2010convex}.

\sidenote{A major advantage of the convex Cayley method is that it is completely unaffected when $\delta$ are intervals rather than exact values~~\cite{sitharam2010convex}.}
%cite my paper with Heping!!!

\myexample
[Using Cayley configuration space].
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic graphs, i.e, its DR-plan has a root and 9 children corresponding to the 9 edges. Solving or recombining the system $(G,\delta)$ corresponding to the root of this DR-plan involves solving a quadratic system with 8 equations and variables. Instead of simultaneously solving this system, we could instead use the fact that $G=K_{3,3}$ is in OMD$_2$: remove the edges $e_1,e_2$ in Figure \ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the non-edges $f_1,f_2$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single interval of attainable length values $\lambda_F$ for the edge $f$. When $\delta$ is generic, i.e, does not admit collinearities or coincidences in the realizations of $(G,\delta)$, the realization space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16 solutions $q^p_{\lambda_f}$ (modulo orientation preserving isometries), with distinct orientation types $\sigma_p$ (two orientation choices for each of the 4 triangles) that can be obtained by solving a sequence of 4 single quadratics in 1 variable (DR-plan of size 2). By subdivided binary search in the interval $\lambda_f \in \Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of $(G,\delta)$ is found when the length of the nonedge $e$ in  the realization $q^p_{\lambda_f}$ is $\delta_e$.

In fact, we can show that $G=K_{3,3}$ is in OMD$_1$ by removing a single edge to reduce (as shown in Figure \ref{fig:omd_k33_example}) to a tree-decomposable graph of low Cayley complexity (which includes the class of partial 2-trees). In the next Section \ref{sec:tdecomp}, we discuss this issue of why the largest class of reduced graphs is desirable.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e, the number of edges to be removed to ensure an isostatic completion that is decomposable with a small DR-plan) by considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than partial 2-trees but nevertheless have convex Cayley configuration spaces at least when the realization space is restricted to a sufficiently comprehensive orientation type. In particular, the so-called \dfn{tree-decomposable graphs of low Cayley complexity} \cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not partial 2-trees. See an illustration in Figure \ref{fig:omd_k33_example}. These too result in DR-plans of size 2 or 3, putting $G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met---for example when $k=1$---because 1-dof Cayley configuration spaces of linkages based on such graphs $G'$ are known to be single intervals when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given. In addition, the bounds of these intervals are of low algebraic complexity. More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3, i.e, the computation of these bounds is tree-decomposable. Alternatively, the bounds are in a simple quadratic or radically solvable extension field over the rationals, or can be computed by solving a triangularized system of quadratics.