### 2015-02-21 doc/GMP2015/revision_1/img/tikz/demo_graph_candrp_seq.tikz.tex

\begin{tikzpicture}
    \tikzstyle{v}=[draw, circle, minimum size=0.75cm]
    \tikzstyle{c}=[draw, circle, inner sep=1.5, fill=black]
    \tikzstyle{b}=[align=center]
    \tikzstyle{e}=[]

    \node[regular polygon, regular polygon sides=3, fill=white, inner sep=4, dashed] (center) at (0,0-.125-.125) {};

% LEVEL 1
    \node[b] (vG) at ( 0,  0) {$G_{\Delta}$};

% LEVEL 2
    \def\x{0}
    \def\y{0}
    % \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    % \tedge{n1va}{n1vb}{solid}{}{};
    \tedge{n1vb}{n1vc}{solid}{}{};
    % \tedge{n1vc}{n1va}{solid}{}{};
    \tedge{n1vb}{n1vd}{solid}{}{};
    \tedge{n1vd}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vb}{solid}{}{};
    \tedge{n1vc}{n1vf}{solid}{}{};
    \tedge{n1vf}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vc}{solid}{}{};

    \def\x{2.5}
    \def\y{-.866}
    \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \tedge{n1va}{n1vb}{solid}{}{};

    \def\x{3.5}
    \def\y{-.866}
    \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    \tedge{n1vc}{n1va}{solid}{}{};

% LEVEL 3
    \def\x{0}
    \def\y{.5-2.35}
    % \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    % \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    % \tedge{n1va}{n1vb}{solid}{}{};
    \tedge{n1vb}{n1vc}{solid}{}{};
    % \tedge{n1vc}{n1va}{solid}{}{};
    \tedge{n1vb}{n1vd}{solid}{}{};
    \tedge{n1vd}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vb}{solid}{}{};
    % \tedge{n1vc}{n1vf}{solid}{}{};
    % \tedge{n1vf}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vc}{solid}{}{};

    \def\x{2}
    \def\y{.5-2.35}
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    \tedge{n1vf}{n1ve}{solid}{}{};

    \def\x{3}
    \def\y{.5-2.35}
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    \tedge{n1vc}{n1vf}{solid}{}{};

% LEVEL 4
    \def\x{0}
    \def\y{.5-4.3}
    % \node[c] (n1va) at (\x+0, \y+0)     [label={left, inner sep=.555}:$a$]{};
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \node[c] (n1vc) at (\x+.5,\y-.866)  [label={right,inner sep=.555}:$c$]{};
    % \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    % \node[c] (n1vf) at (\x+1, \y-1.732) [label={right,inner sep=.555}:$f$]{};
    % \tedge{n1va}{n1vb}{solid}{}{};
    \tedge{n1vb}{n1vc}{solid}{}{};
    % \tedge{n1vc}{n1va}{solid}{}{};
    % \tedge{n1vb}{n1vd}{solid}{}{};
    % \tedge{n1vd}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vb}{solid}{}{};
    % \tedge{n1vc}{n1vf}{solid}{}{};
    % \tedge{n1vf}{n1ve}{solid}{}{};
    \tedge{n1ve}{n1vc}{solid}{}{};

    \def\x{3}
    \def\y{.5-4.3}
    \node[c] (n1vb) at (\x-.5,\y-.866)  [label={left, inner sep=.555}:$b$]{};
    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \tedge{n1vb}{n1vd}{solid}{}{};

    \def\x{4}
    \def\y{.5-4.3}
    \node[c] (n1vd) at (\x-1, \y-1.732) [label={left, inner sep=.555}:$d$]{};
    \node[c] (n1ve) at (\x+0, \y-1.732) [label={below,inner sep=.555}:$e$]{};
    \tedge{n1vd}{n1ve}{solid}{}{};

% DRP edges
    \draw (vG)--(0,  -.75);
    \draw (vG)--(2.1,-.65);
    \draw (vG)--(3.2,-.55);

    \draw (0,  -.2-1.85)--(0,  -.75-1.85);
    \draw (0,  -.2-1.85)--(2.1,-.65-1.85);
    \draw (0,  -.2-1.85)--(3.2,-.55-1.85);

    \draw (0,  -.2-1.85-1.85)--(0,  -.75-1.85-1.85);
    \draw (0,  -.2-1.85-1.85)--(2.1,-.65-1.85-1.85);
    \draw (0,  -.2-1.85-1.85)--(3.2,-.55-1.85-1.85);

\end{tikzpicture}