### 2015-05-07 doc/GMP2015/revision_2/appendix_defs.tex
Major changes to paper (revision 2)

\section{Basic Definitions and Theory}
\label{sec:appendix:defs}
% \label{sec:prelim:defs}

Moved into prelims section.