### 2015-05-16 doc/GMP2015/revision_2/response_to_reviewers.txt
Revision 2. 24 hours to submission.

Response to Reviewers
=====================

Reviewer #1
-----------
Overall I think this paper is timely and interesting.  It is very nice that the authors have included a number of open questions to emphasize that there is more work to be done in this area.  The notation and terminology is rather dense, however, this is probably unavoidable given the intricacy of the questions being addressed.

I don't think the abstract is significantly improved from the original version; there is still a lot of area-specific terminology that makes it hard to understand what the paper is about.  Perhaps a simple motivating example explained in detail over a couple sentences would help, especially explaining why one would want to model a particular ``quasi-uniform, self-similar, layered material structure'' using a decomposition of geometric constraint systems.

To that end, I would also suggest bringing the definition of geometric constraint system out of Appendix A and putting it right at the beginning of section 2, or perhaps section 2.1.

This line at the end of the abstract and at the end of section 2.1 should be modified for the final version of the paper (and applicable files should be included accordingly): ``We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.''

This paper seems essentially ready for publication.

Response:
- Abstract had substantial rework. While we don't give a complete example in the abstract, we do mention which materials we model in this paper. In the introduction, we do explain much more clearly why we would want to use qusecs.
- We moved the definitions into the body of the paper.
- We provide a link where the video and code will be made available. Currently a work in progress, but we are committed to providing a well-documented, open source, highly functional application. The video will be made sooner.

Reviewer #2
-----------
This paper proposes a recursive decomposition planning algorithm to model a material layer as a solution of a 2D geometric constraint system. While the work seems important in material modeling, the topic is a kind of out of focus of GMP audiences. Besides, I'm not quite satisfied with the organization and structure of the paper, and I feel it is not suitable for publication in current form.

1. The paper is far out of a 10 pages limit.
2. The basis concepts should be better put in the main context, but was put in an appendix (Appendix A).
3. The main algorithm in Section 4 should be stood out seperately and explained in details.
4. The examples should be enumerated.
5. The texts contain various formats and styles: Captioned, Bold fonts, colored, underlined, .... This unusual in scientific papers.

Response:
1. We are now submitting to the main journal.
2. We moved preliminaries into the main body of the paper (section 2).
3. We moved the algorithm into its own subsection and revised the proof to be more clear.
4. ...
5. We have addressed this. We have reformatted our unusual headings and such.

Further Changes
===============

Here we discuss additional changes the reviewers should be aware of.

One of the main issues with our paper was that it would not be a good fit for GMP. After much discussion, we decided that we agree and are now submitting the paper for consideration to be published in the standard journal (see next section in this file). The changes made in this revision are discussed below.

Title: Because of the substantial changes to the paper, we decided the title no longer made sense in the context of the revised paper. As such, we have changed it to `Optimal Decomposition and Recombination of Isostatic Geometric Constraint Systems for Designing Layered Materials'. We believe this also makes the intent of our paper much more clear.

Throughout: Many minor corrections to spelling, grammar, and punctuation. Additionally, many sentences were rephrased/restructured to improve clarity. Also, because our paper is no longer being considered for the special issue and does not have to adhere to the page limit, we have reformatted parts of the paper to use more space to improve readability and conform to standard formatting (addressing concerns of reviewer 2).

Two concepts were further fleshed out and discussed in this iteration: (a) The comparison of our canonical DRP to the previously studied cluster-minimal DRP, and (b) the extension of our theory to a larger class of (hyper)graphs, formally those with an underlying abstract rigidity matroid in which independence corresponds to some type of sparsity, and maximal independence (rigidity) is a tightness condition.. This is reflected throughout the paper, but is especially prevalent in (1), (2), (3), and (8).

Abstract: Substantially altered for clarity. Similar content.

(1) Introduction: Entirely rewritten. A major complaint was that it was difficult to get a big picture understanding of the paper without reading the entire thing. We believe we addressed this with our revised intro (and abstract). This also includes a thorough explanation of `qusecs' and why we believe this notion is important to define and give a name, directly addressing reviewer criticism. Also, the discussion of previous work was moved into the corresponding sections (2) and (4).

(2) Preliminaries: Moved back in the definitions from the appendix (as per reviewer request). Moved in discussion of previous work on DR-plans from (1).

(3) Canonical DR-plan: Improved the section introduction. Clarified/improved the definition of a canonical DR-plan. There are new figures to demonstrate the concepts; instead of the previous trivial (in the sense that it was tree-decomposable) example of three triangles, we use three $C2\times C3$s intersecting on a single triangle. Also, there is a new figure that corrects the previous, erroneous figure discussing cluster minimality in the previous revision. Revised proofs for clarity, but there were no fundamental changes. Put algorithm in a separate section (improved proof as mentioned) as per reviewer 2's request. Moved out cluster minimality conjecture to the conclusion. Added a discussion on the overconstrained problem and NP-hardness (3.3).

(4) Recombination: Improved the section introduction. Moved in discussion of previous work on recombination from (1). Made the section that followed in the previous revision into a subsection (4.5).

(5) and (6) Application sections: No changes.

(8) Conclusion: Moved in conjectures about cluster minimality from (3). Added new open problem (OP2).

(App. A) Proofs: Revised DRP proofs for clarity, but there were no fundamental changes.

Journal Change
==============

Originally the paper was submitted for consideration of publication in the special issue of CAGD for GMP2015. However, pending our major revision, this was changed to the regular issue of CAGD. We received confirmation from the GMP co-chairs that this paper would now be handled by the CAGD editors-in-chief. Below, we have copied the email we received from the conference chairs, stating the change.

Upon submission, the exact content of this section will also be emailed to editors-in-chief, along with our submission number: XX.

If there are additional questions in regard to this issue, please send an email.

Email
-----
GMP2015 <gmp2015@easychair.org> Wed, Mar 25, 2015 at 7:17 AM
To: Troy Baker <tbaker@cise.ufl.edu>
Dear Troy Baker,

even though your GMP submission will not appear in the conference special
issue, but - after a major revision - in a forthcoming regular issue of
CAGD, you still have the chance to present your work at the GMP 2015
conference, and we would be very glad if you accept this invitation.

More information about the conference can be found here:
http://gmp2015.inf.usi.ch/

Note that you may require a visa to come to Switzerland and that the
application process can take some time. So it is best to start early. More
information about visa can be found here: http://gmp2015.inf.usi.ch/visa

Please inform us about your decision whether or not to come to GMP 2015
and to present your work by sending an e-mail to "gmp2015@usi.ch" before
*April 2, 2015* to help us plan the conference program.

With kind regards

Elaine Cohen & Kai Hormann
[GMP 2015 conference chairs]